---
title: "arima_garchv2"
author: "Moinak Nath"
date: "2025-03-30"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. Getting Stock Data

```{r}
library(quantmod)
library(PerformanceAnalytics)
```

```{r}
# Define the ticker symbol as a variable
ticker <- "AAPL"

# Define the date range
start_date <- "2016-04-08"
end_date   <- "2025-04-10"

# Get data using the ticker variable
data <- getSymbols(ticker, src = "yahoo", from = start_date, to = end_date, auto.assign = FALSE)
close_prices <- Cl(data)
returns <- dailyReturn(close_prices)


```

```{r}
cat("Number of days: ", length(close_prices))

```

## Basic Plots

```{r}

# Plot closing prices with dynamic title using sprintf
plot(close_prices, main = sprintf("%s Close", ticker))

# Plot daily returns with dynamic title using sprintf
plot(returns, main = sprintf("%s Daily Returns", ticker), col = "blue", type = "l")

```

```{r}
library(forecast)
library(rugarch)

arima_garch_prediction <- function(time_series, initial_window, days_ahead) {
  n <- length(time_series)
  forecasted_prices <- list()
  forecasted_volatilities <- list()
  
  for (i in seq(initial_window, n - days_ahead)) {
    training_data <- time_series[1:i]
    
    # Fit ARIMA
    fit <- auto.arima(training_data)
    arima_res <- residuals(fit)
    
    # GARCH(1,1) specification
    spec <- ugarchspec(
      variance.model = list(model = "sGARCH", garchOrder = c(1, 1)),
      mean.model     = list(armaOrder = c(0, 0), include.mean = FALSE),
      distribution.model = "norm"
    )
    
    # Fit GARCH to residuals
    garch_fit <- tryCatch(
      ugarchfit(spec = spec, data = arima_res, solver.control = list(trace = 0)),
      error = function(e) NULL
    )
    
    if (is.null(garch_fit)) {
      warning(paste("GARCH model failed to fit at index", i))
      next
    }
    
    # Forecast volatility
    garch_forecast <- ugarchforecast(garch_fit, n.ahead = days_ahead)
    forecasted_volatility <- as.numeric(garch_forecast@forecast$sigmaFor[1, ])
    
    # Forecast mean from ARIMA
    arima_forecast <- forecast(fit, h = days_ahead)
    arima_point_forecast <- as.numeric(arima_forecast$mean)
    
    # Simulate prices with random noise
    random_shocks <- forecasted_volatility * rnorm(days_ahead)
    final_forecasted_prices <- arima_point_forecast + random_shocks
    
    # Store results
    forecasted_prices[[length(forecasted_prices) + 1]] <- final_forecasted_prices
    forecasted_volatilities[[length(forecasted_volatilities) + 1]] <- forecasted_volatility
  }
  
  return(list(
    prices = forecasted_prices,
    volatilities = forecasted_volatilities
  ))
}

```

```{r}
print(length(log(close_prices)))
res = arima_garch_prediction(log(close_prices), 10, 1)
prices = res[1]
vol = res[2]
```

```{r}
x = prices$prices[1517]
print(prices$prices[157])
```

```{r}
library(quantmod)
library(PerformanceAnalytics)
library(forecast)
library(rugarch)

log_prices <- log(close_prices)

# Run ARIMA-GARCH predictions
prices <- unlist(res$prices)
vol <- unlist(res$volatilities)

# Create aligned actual log prices
actual_log <- as.numeric(log_prices)
forecast_start_index <- 11  # Because initial_window = 5
actual_log_aligned <- actual_log[forecast_start_index:(forecast_start_index + length(prices) - 1)]
plot(actual_log_aligned, type = "l", col = "black", lwd = 2,
     main = sprintf("Forecasted vs Actual Log(%s) Prices", ticker),
     ylab = "Log Price", xlab = "Days")
# 📈 Plot: Forecasted vs Actual Log Prices
plot(actual_log_aligned, type = "l", col = "black", lwd = 2,
     main = sprintf("Forecasted vs Actual Log(%s) Prices", ticker),
     ylab = "Log Price", xlab = "Days")
lines(prices, col = "blue", lwd = 2)
legend("topleft", legend = c("Actual", "Forecasted"), col = c("black", "blue"), lwd = 2)


```

```{r}
library(zoo)

# Compute daily log returns
log_returns <- diff(as.numeric(log_prices))

# Define rolling window size for realized volatility (e.g., 5 days)
window_size <- 5

# Compute rolling standard deviation (realized volatility)
realized_vol <- rollapply(log_returns, width = window_size, FUN = sd, align = "right", fill = NA)

# Align with forecasted volatility
# The forecasts start from index 6, so realized_vol must be trimmed accordingly
start_index <- 10 + 1  # initial_window + 1 (because of diff)
realized_vol_aligned <- realized_vol[start_index:(start_index + length(vol) - 1)]

# 🟦 Plot forecasted vs realized volatility
plot(realized_vol_aligned, type = "l", col = "black", lwd = 2,
     main = paste0("Forecasted vs Realized Volatility (", window_size, "-day rolling SD)"),
     ylab = "Volatility", xlab = "Days")
lines(vol, col = "red", lwd = 2)
legend("topleft", legend = c("Realized", "Forecasted"), col = c("black", "red"), lwd = 2)

```

## Longer Forecast

```{r}
arima_garch_prediction_long_forecast <- function(time_series, initial_window, days_ahead) {
  n <- length(time_series)
  forecasted_prices <- rep(NA, n)        # Pre-fill with NAs
  forecasted_volatilities <- rep(NA, n)  # Same length as time_series

  # Step forward in chunks of 'days_ahead'
  for (i in seq(initial_window, n - days_ahead, by = days_ahead)) {
    training_data <- time_series[1:i]
    
    # Fit ARIMA
    fit <- auto.arima(training_data)
    arima_res <- residuals(fit)
    
    # GARCH(1,1) specification
    spec <- ugarchspec(
      variance.model = list(model = "sGARCH", garchOrder = c(1, 1)),
      mean.model     = list(armaOrder = c(0, 0), include.mean = FALSE),
      distribution.model = "norm"
    )
    
    # Fit GARCH to residuals
    garch_fit <- tryCatch(
      ugarchfit(spec = spec, data = arima_res, solver.control = list(trace = 0)),
      error = function(e) NULL
    )
    
    if (is.null(garch_fit)) {
      warning(paste("GARCH model failed to fit at index", i))
      next
    }
    
    # Forecast volatility
    garch_forecast <- ugarchforecast(garch_fit, n.ahead = days_ahead)
    forecasted_volatility <- as.numeric(garch_forecast@forecast$sigmaFor[1, ])
    
    # Forecast price via ARIMA
    arima_forecast <- forecast(fit, h = days_ahead)
    arima_point_forecast <- as.numeric(arima_forecast$mean)
    
    # Simulate price path
    random_shocks <- forecasted_volatility * rnorm(days_ahead)
    final_forecasted_prices <- arima_point_forecast + random_shocks
    lower_bound <- arima_point_forecast - 1.96 * forecasted_volatility
    upper_bound <- arima_point_forecast + 1.96 * forecasted_volatility

    
    # Store forecasts
    indices <- (i + 1):(i + days_ahead)
    if (max(indices) <= n) {
      forecasted_prices[indices] <- final_forecasted_prices
      forecasted_volatilities[indices] <- forecasted_volatility
    }
  }

  return(list(
    prices = forecasted_prices,
    volatilities = forecasted_volatilities,
    low = lower_bound,
    high = upper_bound
  ))
}

```

```{r}
res1 = arima_garch_prediction_long_forecast(log(close_prices), 10, 5)
prices1 = res1[1]
vol1 = res1[2]
low1 = res1[3]
high1 = res1[4]
```

```{r}
str(prices1)
print(exp(prices1$prices[2265]))

```

```{r}
str(low1)
```

```{r}

library(quantmod)
library(PerformanceAnalytics)
library(forecast)
library(rugarch)

log_prices <- log(close_prices)

# Run ARIMA-GARCH predictions
prices <- unlist(res1$prices)
vol <- unlist(res1$volatilities)

# Create aligned actual log prices
actual_log <- as.numeric(log_prices)
forecast_start_index <- 11  # Because initial_window = 5
actual_log_aligned <- actual_log[forecast_start_index:(forecast_start_index + length(prices) - 10 - 1)]

# Shift the forecasted series left by one step to align with actual data
shift_left <- function(x, k = 1) {
  c(x[(k + 1):length(x)], rep(NA, k))
}

cat("Length of prices: ", length(prices), "\n")
cat("Length of QQQ returns: ", length(actual_log_aligned), "\n")

print(prices)
# Example: align price forecasts
forecasted_prices_aligned <- prices[forecast_start_index:(forecast_start_index + length(prices) - 10 - 1)]

plot(actual_log_aligned, type = "l", col = "black", lwd = 2,
     main = "Forecasted vs Actual Log(QQQ) Prices",
     ylab = "Log Price", xlab = "Days")
# 📈 Plot: Forecasted vs Actual Log Prices
plot(actual_log_aligned, type = "l", col = "black", lwd = 2,
     main = "Forecasted vs Actual Log(QQQ) Prices",
     ylab = "Log Price", xlab = "Days")
lines(forecasted_prices_aligned, col = "blue", lwd = 2)
legend("topleft", legend = c("Actual", "Forecasted"), col = c("black", "blue"), lwd = 2)


```

```{r}

library(zoo)

# Compute daily log returns
log_returns <- diff(as.numeric(log_prices))

# Define rolling window size for realized volatility (e.g., 5 days)
window_size <- 10

# Compute rolling standard deviation (realized volatility)
realized_vol <- rollapply(log_returns, width = window_size, FUN = sd, align = "right", fill = NA)

# Align with forecasted volatility
# The forecasts start from index 6, so realized_vol must be trimmed accordingly
start_index <- 10 + 1  # initial_window + 1 (because of diff)
realized_vol_aligned <- realized_vol[start_index:(start_index + length(vol) - 1)]
forecasted_vol_aligned <- shift_left(vol, k = 10)

# 🟦 Plot forecasted vs realized volatility
plot(realized_vol_aligned, type = "l", col = "black", lwd = 2,
     main = paste0("Forecasted vs Realized Volatility (", window_size, "-day rolling SD)"),
     ylab = "Volatility", xlab = "Days")
lines(forecasted_vol_aligned, col = "red", lwd = 2)
legend("topleft", legend = c("Realized", "Forecasted"), col = c("black", "red"), lwd = 2)

```

## TO-DO

-   Clean up code

-   Metrics: Mean squared error, what percent of time did it predict the direction of price movement for next days_ahead days, percent error in volatility daily

-   bands

-   Generate strategies: Sharpe ratio, buy if predict up, sell if predict down, hedging

    -   Delta hedging: for short time period estimates the volatility forecast as a kind of average, so if it surges above that hedge and bet that it will fall to the mean
